---
title: "Getting started with connector.databricks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{connectordatabricks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The connector.databricks package provides a convenient interface for accessing and 
interacting with Databricks tables and volumes directly from R. 


## Introduction 
This vignette will guide you through the process of connecting to a Databricks, 
retrieving data, and performing various operations using this package.

## Connecting to a Databricks
### Without connector package

To get started, you need to establish a connection to your Databricks cluster
or volume storage. Use:

* `connector_databricks_table()` function to authenticate and connect to your Databricks cluster
* `connector_databricks_volume()` function to connect to your Databricks volume storage

Here's an example of how to do this:

```{r, eval=FALSE}
library(connector.databricks)

# Connect to databricks tables using DBI
con <- connector_databricks_table(
  httpPath = "path-to-cluster",
  catalog = "my_catalog",
  schema = "my_schema"
)

# Connect to databricks volume
con <- connector_databricks_volume(
  catalog = "my_catalog",
  schema = "my_schema",
  path = "path-to-file-storage"
)
```

### With connector package (recommended)

If you are using the connector package, you can connect to a Databricks datasources 
using the `connect()` function. This function based on a configuration file or 
a list creates a `connectors()` object with a `connector`for each of the 
specified datasources (for detailed explanation have a look at the `connector` package). 

Configuration file for connecting to Databricks should look like this:

```{yaml}
# nolint start
metadata:
  catalog:  "databricks_calatog"
  http_path: "path-to-cluster"
  path: "path-to-file-storage"
  project: "project_name"
  trial: "trial_name"

datasources:
  - name: "tables"
    backend:
      type: "connector.databricks::connector_databricks_table"
      http_path: "{metadata.http_path}"
      catalog: "{metadata.catalog}"
      schema: "{metadata.project}_{metadata.trial}_adam"
  - name: "volume"
    backend:
      type: "connector.databricks::connector_databricks_volume"
      path: "{metadata.path}"
      catalog: "{metadata.catalog}"
      schema: "{metadata.project}_{metadata.trial}_tfl"
# nolint end
```

Save this to `_connector.yml` file and use the `connect()` function to connect 
to Databricks:

```{r, eval=FALSE}
library(connector)

# Create connector object
db <- connect()
```

Now you can access the Databricks cluster and volume using the `db` object and 
`tables` or `volume` field, respectively.


```{r, eval=FALSE}
# Connection to Databricks cluster. This will print object details
db$tables

# Connection to Databricks cluster. This will print object details
db$volume
```

After the setup is done we can use these connections to manipulate Databricks data.

## Listing data

You can do some basic directory operations, such as creating, removing, and listing
data inside directory, or listing tables inside database.

```{r, eval=FALSE}
# Create a directory
db$volume |>
  create_directory_cnt("new_directory")

# Remove a directory
db$volume |>
  remove_directory_cnt("new_directory")

# List content inside volume directory
db$volume |>
  list_content_cnt()

# List tables inside database
db$tables |>
  list_content_cnt()
```


## Reading and writing data

The `connector` packages provide a set of functions to read and write data from/to
the datasources. They all have similar interface, so it's easy to switch between
them.

Now, we will show how to read and write different types of data from/to Databricks.
In these examples we will be using `iris` and `mtcars` datasets.

Here is an example of writing data to a table on a cluster:

```{r eval=FALSE}
library(dplyr)

# Manipulate data

## Iris data
setosa <- iris |>
  filter(Species == "setosa")

mean_for_all_iris <- iris |>
  group_by(Species) |>
  summarise_all(list(mean, median, sd, min, max))

## Mtcars data
cars <- mtcars |>
  filter(mpg > 22)

mean_for_all_mtcars <- mtcars |>
  group_by(gear) |>
  summarise(across(
    everything(),
    list(
      "mean" = mean,
      "median" = median,
      "sd" = sd,
      "min" = min,
      "max" = max
    ),
    .names = "{.col}_{.fn}"
  )) |>
  tidyr::pivot_longer(
    cols = -gear,
    names_to = c(".value", "stat"),
    names_sep = "_"
  )

## Store data
db$tables |>
  write_cnt(setosa, "setosa", overwrite = TRUE)

db$tables |>
  write_cnt(mean_for_all_iris, "mean_iris", overwrite = TRUE)

db$tables |>
  write_cnt(cars, "cars_mpg", overwrite = TRUE)

db$tables |>
  write_cnt(mean_for_all_mtcars, "mean_mtcars", overwrite = TRUE)
```

Now, let's read the data back manipulate it a bit and write it to the Databricks
volume. This way we can save different types of data in different formats.

```{r eval=FALSE}
library(gt)
library(tidyr)
library(ggplot2)

# List and load data from cluster
db$tables |>
  list_content_cnt()

table <- db$tables |>
  read_cnt("mean_mtcars")

gttable <- table |>
  gt(groupname_col = "gear")

# Save nontabular data to databricks volume
tmp_file <- tempfile(fileext = ".docx")
gtsave(gttable, tmp_file)
db$volume |>
  upload_cnt(tmp_file, "tmeanallmtcars.docx")

# Manipulate data
setosa_fsetosa <- db$tables |>
  read_cnt("setosa") |>
  filter(Sepal.Length > 5)

fsetosa <- ggplot(setosa) +
  aes(x = Sepal.Length, y = Sepal.Width) +
  geom_point()

## Store data into output location
db$volume |>
  write_cnt(fsetosa$data, "fsetosa.csv")
db$volume |>
  write_cnt(fsetosa, "fsetosa.rds")

tmp_file <- tempfile(fileext = ".png")
ggsave(tmp_file, fsetosa)
db$volume |>
  upload_cnt(tmp_file, "fsetosa.png")
```

## Conclusion

In this vignette we showed how to connect to Databricks datasources, read and write
data from them. We also showed how to use the connector package to connect to
Databricks and how to manipulate data using the connector package.
